---
title: "AGEC 840 Tariff Model"
author: "Nelson Villoria"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AGEC 840 Tariff Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

This vignette demonstrates the core workflow for the `agec840tariffmodel`
package. The package implements a one-good partial equilibrium model with
constant elasticities, calibrated to a benchmark equilibrium and used to
analyze tariffs, production subsidies, consumption subsidies, and quotas.

## Calibration

We start by calibrating the model to benchmark data.

```{r calibration}
library(agec840tariffmodel)
cal <- calibrate.model(
  kd     = 1000,
  ks     = 700,
  md     = 300,
  eta.d  = -0.30,
  eps.s  = 0.50,
  eps.ms = 5.00,
  tau    = 0.10
)
cal
```

## Baseline equilibrium

Solve the benchmark equilibrium using the calibrated model.

```{r baseline}
base <- solve.model(cal)
print.solution(base, "Baseline: 10% tariff")
```

## Tariff counterfactual

Raise the tariff to 20% and compare the counterfactual to the baseline.

```{r tariff}
cf <- solve.model(cal, tau = 0.20)
print.solution(cf, "Counterfactual: 20% tariff")
wf <- welfare.change(base, cf, cal)
print.welfare(wf, "Welfare change: 10% -> 20% tariff")
```

## Quota example

Run a quota counterfactual with domestic rent capture.

```{r quota}
cf.quota <- solve.model(cal, tau = 0, quota = 240, rent.domestic = 1)
print.solution(cf.quota, "Quota: 240 units, domestic rents")
wf.quota <- welfare.change(base, cf.quota, cal)
print.welfare(wf.quota, "Welfare change: baseline -> quota")
```

## Price transmission

The package also includes a log-linear approximation for the price
transmission elasticity.

```{r transmission}
approx <- linear.approx(cal, tau = 0.20)
approx
```
